---
title: "Searching for Images from a DPO Mass Digitization Project"
author: "Luis J. Villanueva"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Searching for Images from a DPO Mass Digitization Project}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

OCIO's Digitization Program Office completed a Mass Digitization project of the Living Orchid Collection of Smithsonian Gardens. As part of a new effort to showcase our projects, we used EDANr to search for the images taken during this project. When a plant had flowers, a close-up photograph of the flower was taken. These close-ups were downloaded using EDANr to display in 4k resolution displays at the DPO office. This vignette summarizes the process:

1. Query EDAN for the specimens in the collection
1. Look for images taken by the contractor in the project
1. Download the close-up images
1. Resize the images to 4k and add text with general information to the images

Article in [Smithsonian Insider](https://insider.si.edu/2018/08/see-thousands-of-orchids-in-incredible-detail-in-the-smithsonians-newly-digitized-collection/) about the digitization project. 

## Query EDAN for the specimens in the collection

Add the AppID and AppKey, load packages, other preliminary steps:

```{r, eval=FALSE}
#EDAN creds
AppID = "APP_ID"
AppKey = "verylong_key"

#Required packages
library("EDANr")
library("magick")
library("stringr")

#Create folders to store images
#Horizontal images
dir.create("images_h", showWarnings = FALSE)
#Vertical images
dir.create("images_v", showWarnings = FALSE)
```

Run the query, first to get the total number of results EDAN has:

```{r, eval=FALSE}
#EDAN query for orchids of Smithsonian Gardens with images
orchids_query <- "orchid smithsonian gardens&fq=online_media_type:\"Images\""
orchids_fqs <- "online_media_type:\"Images\""


#Get number of results in EDAN
results <- EDANr::edan_metadata_search(query = orchids_query, 
                             fqs = orchids_fqs,
                             AppID = AppID, 
                             AppKey = AppKey, 
                             rows = 1, 
                             start = 0)
results_count <- results$rowCount

#Calculate the number of steps needed to get all the results,
# in steps of 100 rows each, the maximum the API returns
steps <- floor(results_count/100)
```

Then, loop over the results in steps of 100 results. 

```{r, eval=FALSE}
#Loop over each step, getting all the images
for (i in seq(0, steps)){
  #Query to get the next 100 results
  results <- EDANr::edan_metadata_search(query = orchids_query, 
                               fqs = orchids_fqs,
                               AppID = AppID, 
                               AppKey = AppKey, 
                               rows = 100, 
                               start = (i*100))
  
  #If there are results in this step
  if (!is.null(dim(results$rows))){
    
    #Loop over each result in this step
    for (j in seq(1, dim(results$rows)[1])){
      #
      #See code below
      #
    }
  }
}
```

## Look for images taken by the vendor

Get the images of the j-th row:

```{r, eval=FALSE}
#Get images, available from IDS
ids_images <- results$rows$content$descriptiveNonRepeating$online_media$media[j][[1]]

#Filter for images that were taken by the contractor of the project
which_image <- ids_images[stringr::str_which(ids_images$caption, "Creekside Digital"),]

#Select the close-up image of the flowers of the plant, coded with '102' in the filename
which_image_detail <- which_image[stringr::str_which(ids_images$idsId, "102"),]
```

## Download the close-up images

Get the URL of the images and download:

```{r, eval=FALSE}
#Get image URL
image <- which_image_detail$content[1]

#Get image ID
idsID <- which_image_detail$idsId[1]

#Download the image
localimage <- paste0(idsID, ".jpg")
dlfile <- try(download.file(url = image, 
                            destfile = localimage, 
                            mode = "wb"), 
              silent = TRUE)

#If there is an error, skip and
# go to the next one in the loop
if (class(dlfile) == "try-error"){
    next
  }
```

## Resize and add text with general information to the images

Resize the image for 4k display, checking if horizontal or vertical first:

```{r, eval=FALSE}
#Read image to find size
img <- image_read(localimage)
img_info <- image_info(img)

if (img_info$width > img_info$height){
  #Horizontal, resize to 3840x2160
  img_scaled <- image_scale(img, "3840x2160")
}else{
  #Vertical, resize to 2160x3840
  img_scaled <- image_scale(img, "2160x3840")
}
```

Get specimen info to write on the image:

```{r, eval=FALSE}
#Get image title, which contains the name of the specimen
img_title <- results$rows$title[j]

#If title is too long, break into two lines
if (nchar(img_title) > 30){
  img_title1 <- strsplit(img_title, " ")[[1]]
  spl_no <- floor(length(img_title1)/2)
  
  img_title_1 <- paste(img_title1[1:spl_no], 
                       collapse = " ")
  img_title_2 <- paste(img_title1[spl_no+1:(length(img_title1) - spl_no)], 
                       collapse = " ")
  img_title <- paste0(img_title_1, 
                      "\n   ", 
                      img_title_2)
}

#Get common name of plant, if available
common_name <- unlist(results$rows$content$indexedStructured$common_name[j])[[1]]
if (!is.null(common_name)){
  title <- paste0(common_name, "\n", img_title)
}

#Add the accession number and identify the unit
title <- paste0(title, 
                "\nAccession Number: ", 
                results$rows$content$freetext$identifier[j][[1]]$content, 
                "\nSmithsonian Gardens")

#Write the info to the image
img_ready <- image_annotate(img_scaled, 
                            title, 
                            font = 'Arial', 
                            size = 60, 
                            color = "white", 
                            gravity = "SouthWest")

#Write image to one of two folders
if (img_info$width > img_info$height){
  #Image is wide, store in Horizontal folder
  image_write(img_ready, 
              path = paste0("images_h/" , 
                            idsID, 
                            ".jpg"), 
              format = "jpg")
}else{
  #Image is tall, store in Vertical folder
  image_write(img_ready, 
              path = paste0("images_v/",
                            idsID, 
                            ".jpg"), 
              format = "jpg")
}

#Delete original image
file.remove(localimage)
```
